---
title: "example_pipeline"
output: html_document
---

```{r}
library(brms)
source("../../bsae/data_wrangling.R")
```

```{r}
files = list.files("../R/")
for (file in files) {
  source(paste0("../R/", file))
}

rm(file, files)
```

# Automatic BHF
```{r}
spec_BHF = specify(
  model = "BHF",
  domain = "county",
  response = "obs_biomass",
  auxiliary_variables = c("tcc", "evc"),
  variable_transform = list(
    transform = function (y) {y^(1/3)},
    inv_transform = function (y) {y^3}
  )
)

fit_BHF = fit.basal_spec(
  spec_BHF, 
  (idaho_obs_full %>% filter(obs_biomass > 0))
)

check_BHF = check.basal_fit(
  fit = fit_BHF
)

estimates_BHF = estimate.basal_fit(
  fit = fit_BHF,
  newdata = idaho_aux_aggregate,
  domain = "county",
  ndraws = 5000
); estimates_BHF$preds
```

# Manual BHF
```{r}
spec_custom = specify(
  formula = formula(obs_biomass ~ tcc + evc + (1 | county)),
  level = "unit",
  variable_transform = list(
    transform = function (y) {y^(1/3)},
    inv_transform = function (y) {y^3}
  )
)

fit_custom = fit.basal_spec(
  spec_custom,
  (idaho_obs_full %>% filter(obs_biomass > 0))
)

check_custom = check.basal_fit(
  fit = fit_custom
)

estimates_custom = estimate.basal_fit(
  fit = fit_custom,
  newdata = idaho_aux_aggregate,
  domain = "county",
  ndraws = 5000
); estimates_custom$preds
```

# Automatic FH
```{r}
spec_FH = specify(
  model = "FH",
  domain = "county",
  response = "obs_biomass",
  auxiliary_variables = c("tcc", "evc"),
  variable_transform = list(
    transform = function (y) {y^(1/3)},
    inv_transform = function (y) {y^3}
  )
)

fit_FH = fit.basal_spec(
  spec_FH,
  idaho_obs_full %>% filter(obs_biomass > 0),
  population_size = nrow(idaho_aux_full)
)

check_FH = check.basal_fit(
  fit = fit_FH
)

estimates_FH = estimate.basal_fit(
  fit = fit_FH,
  newdata = idaho_aux_aggregate,
  domain = "county",
  ndraws = 5000
); estimates_FH$preds
```

# Manual FH
```{r}
spec_area_custom = specify(
  formula = formula(obs_biomass ~ tcc + evc + (1 | county)),
  level = "area",
  domain = "county"
)

fit_area_custom = fit.basal_spec(
  spec_area_custom,
  idaho_obs_full %>% filter(obs_biomass > 0),
  population_size = nrow(idaho_aux_full)
)

check_area_custom = check.basal_fit(
  fit = fit_area_custom
)

estimates_area_custom = estimate.basal_fit(
  fit = fit_area_custom,
  newdata = idaho_aux_aggregate,
  ndraws = 5000
); estimates_area_custom$preds
```

# Automatic FH, pre-aggregated

```{r}
spec_FH_pre_agg = specify(
  model = "FH",
  domain = "county",
  response = "obs_biomass",
  auxiliary_variables = c("tcc", "evc"),
  obs_variability = fit_FH$data$BASAL_HT_SE # also works with "BASAL_HT_SE"
)

fit_FH_pre_agg = fit.basal_spec(
  spec_FH_pre_agg,
  fit_FH$data
)

check_FH_pre_agg = check.basal_fit(
  fit = fit_FH_pre_agg
)

estimates_FH_pre_agg = estimate.basal_fit(
  fit = fit_FH_pre_agg,
  newdata = idaho_aux_aggregate,
  domain = "county",
  ndraws = 5000
); estimates_FH$preds
```

```{r}
second_stage <- specify(
  model = "BHF",
  auxiliary_variables = c("tcc", "evc"),
  specifying_second_stage_model = TRUE
)

two_stage_spec <- specify(
  level = "unit",
  model = "BHF",
  response = "obs_biomass",
  auxiliary_variables = "tcc",
  domain = "county",
  model_stage = "zi",
  second_stage_spec = second_stage,
  variable_transform = list(
    transform = function (y) {y^(1/3)},
    inv_transform = function (y) {y^3}
  )
)

two_stage_fit <- fit.basal_spec(
  two_stage_spec,
  idaho_obs_full
)

two_stage_check <- check.basal_fit(
  two_stage_fit
)

estimate_two_stage = estimate.basal_fit(
  two_stage_fit,
  newdata = idaho_aux_full,
  domain = "county",
  max_preds = 1e5,
  ndraws = 200
); estimate_two_stage
```


To do:
1. Allow user to specify priors
2. (maybe) implement generics like `predict()`, `posterior_predict()`, or `post_epred()`
  - would make things like posterior predictive checks for two-stage models and similar things much much easier
  - While we're at it, we could extend `summary()` to `fit.basal` class
3. Implement ZI estimator


Questions:
1. domain vs area for "small areas"


