---
title: "Advanced Specification"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
mean = mean
sd = sd
quantile = quantile
```

```{r setup}
library(basal)
library(tidyverse)

data("oregon_population", "oregon_sample")
```

\section{Data Description}

Here's some data. There will be a better section here, or no section, depending. Either way, these are FIA data like what we might use, but they are public. We have a number of columns, of which I have selected the interesting ones.

```{r}
oregon_population = oregon_population |>
  dplyr::select(c("countyfips", "tnt", "evt_class_cd", "ppt", "elev", "lfevc"))
oregon_sample = oregon_sample |>
  dplyr::select(c("biomass", "countyfips", "tnt", "evt_class_cd", "ppt", "elev", "lfevc"))
```

To assist with stability and reduce variability of priors, we will scale the `elev` variable to have a smaller variance. We will discuss this more in depth in another vignette---`advanced_model_fitting`---but take it for granted here.

```{r}
# sapply(oregon_sample, sd) # find variables with large variances

oregon_population$elev = oregon_population$elev/sd(oregon_sample$elev)
oregon_sample$elev = oregon_sample$elev/sd(oregon_sample$elev)

oregon_population$ppt = oregon_population$ppt/sd(oregon_sample$ppt)
oregon_sample$ppt = oregon_sample$ppt/sd(oregon_sample$ppt)
```

\section{Details of `specify`}
While a model may be specified quite easily, there can be a good amount of nuance in a specification. We will start with a BHF and expand outwards into increasingly complex models. A basic BHF can be specified like below with `model = "BHF"`.

```{r}
bhf <- specify(
  model = "BHF", 
  auxiliary_variables = c("lfevc", "ppt"),
  domain = "evt_class_cd", 
  response = "biomass"
)
```

Although, as alluded to in the vignette on basic workflow, a variable transformation is often useful for acceptable model fitness. This can be specified through supplying a transformation as well as an inverse transformation. The inverse transformation should be a left-inverse of the transformation. The response variable for the model is transformed before model fitting in `fit`, and then before returning values in `estimate`. If estimating a two-stage model, then the response values are transformed before weighting pixels by their probability of nonzero values. \textcolor{red}{I will add more to this once this is more interesting to talk about more.}

```{r}
bhf <- specify(
  model = "BHF", 
  auxiliary_variables = c("lfevc", "ppt"),
  domain = "evt_class_cd", 
  response = "biomass",
  variable_transform = list(
    transform = function (y) {y^(1/4)},
    inv_transform = function (y) {y^4}
  )
)
```





\section{Modeling parameters of the model explicitly}
A common issue with linear regression is heteroscedasticity, or data with different residual variances. Models which try to fix this are possible in the Bayesian context, and quite simple with `brms`, and by extension `basal`.

```{r}
logit_dvi = specify(
  model = "BHF",
  auxiliary_variables = c("lfevc", "ppt"),
  domain = "evt_class_cd",
  specifying_second_stage = TRUE
)

dvi_spec = specify(
  brms::brmsformula(biomass ~ lfevc + ppt + (1 | evt_class_cd),
              sigma ~ (1 | evt_class_cd)),
  level = "unit",
  variable_transform = list(
    transform = function (y) {y^(1/4)},
    inv_transform = function (y) {y^4}
  ),
  model_stage = "zi",
  second_stage_spec = logit_dvi
)

dvi_fit = fit.basal_spec(
  dvi_spec,
  oregon_sample,
  ncores = 3
)

check.basal_fit(
  dvi_fit,
)

estimates <- estimate.basal_fit(
  fit = dvi_fit,
  newdata = oregon_population,
  max_preds = 1e5,
  ndraws = 800
); estimates
```

```{r}
estimates
```


