---
title: "Workflow"
author: Leland Russell, Wystan Wu
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: |
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
library(brms)
library(bayesplot)
source("../../bsae/data_wrangling.R")

files = list.files("../R/")
for (file in files) {
  source(paste0("../R/", file))
}

rm(file, files)
```


\section{Basic Workflow}
The governing philosophy behind `basal` is that every model is, up to specification details, the same. Frequentist models may give the illusion that different types of models are fundamentally different. For example, OLS estimators have a closed form and can be directly estimated. Meanwhile, generalized linear models require iterative methods for parameter estimations or mixed models use restricted maximum likelihood. The different methods for estimating models separate them into different strata, different functions for estimation, different modeling assumptions.

In the Bayesian paradigm, however, MCMC methods ensure that every model is estimated the same way. To reflect this, every regression model listed above, and indeed more not covered within this package, are the same. The defining factor of a model is the specification of parameters and their contribution to a response variable. Although this may be relaxed to existence of certain parameters, as full specification (i.e., prior distribution specification) is not always as necessary.

When modeling in `basal`, every model starts with a specification which will determine, up to prior distributions, the model. These can be simple models often used within the literature, like a Battese-Harter-Fuller (BHF) or Fay-Herriot (FH) models, or they can be more complex custom models. For example, to specify a BHF model, one need only call

```{r}
BHF_spec <- specify(
  model = "BHF",
  domain_name = "county",
  response_name = "obs_biomass",
  auxiliary_variables = c("tcc", "evc")
)
```

From this, the level of the model as well as the relationship between the response and covariates (the auxiliary variables) will be well understood and unambiguous. One could also wish to specify a custom model and create the same model as above:

```{r}
BHF_spec_manual <- specify(
  obs_biomass ~ tcc + evc + (1 | county),
  level = "unit"
)
```

The second specification is both simpler and more complex. It requires less input from the user provided they know how to express their modelling assumptions in a form like that of {lme4}. Simultaneously, though, it may be more ambiguous to express a model by a formula. Either way, these two models will impose the same modeling assumptions and ultimately be indistinguishable.

In the transition to the area-level model, while modeling assumptions may become more stringent, the burden on the user changes only marginally. When estimating a FH model, for example, one may use almost identical code as the BHF model, substituting only `model = "BHF"` for `model = "FH"`.

```{r}
FH_spec <- specify(
  model = "FH",
  domain_name = "county",
  response_name = "obs_biomass",
  auxiliary_variables = c("tcc", "evc")
)
```

From this, the engine will understand that, as the FH is an area-level model, it is estimating the direct estimators, and the variance of these estimators should be known from the estimation procedure. It will compute the direct estimators for the given response, and impose the restriction that the variance of these estimators is known. Accordingly though, it must know what domain to aggregate over. In the case of the named FH model, this is not an issue as the domain name must already be provided, but for a custom model, the user must provide this as well:

```{r}
FH_spec_manual <- specify(
  obs_biomass ~ tcc + evc + (1 | county),
  level = "area",
  domain_name = "county"
)
```

Once a model has been specified, its parameters can be estimated through the `fit()` function. While there are many options for `fit()`, the only necessary ones are a previously specified model, and a dataset. The columns in the dataset must correspond to the formulae or various variable names listed in `specify()`, although aside from area-level models, not much is needed.

```{r, cache = TRUE, include=FALSE}
BHF_fit <- fit.basal_spec(
  BHF_spec,
  idaho_obs_full
)
```

For area level models, in order to compute direct estimates, the size of the population must be given.

```{r, cache = TRUE, include=FALSE}
FH_fit <- fit.basal_spec(
  FH_spec,
  idaho_obs_full,
  population_size = nrow(idaho_aux_full)
)
```

With these models in hand, it is best practices to ensure the model specification and subsequent MCMC posterior approximation is accurate, which we do with the `check()` function.

```{r}
check.basal_fit(BHF_fit)
```

The output from these functions shows statistics computed from the original population (black) as well as statistics computed from posterior predictive populations (blue). The first plow shows the empirical density of both, and the fourth shows the empirical cdf. These are both related and show the same information in different ways. In both, we see a significant deviation of our posterior predictive samples from the observed sample indicating that our model does not explain the data very well. 

The other two plots show the density of the distribution of posterior predictive means and variances with lines indicating the point statistics calculated from the provided data. These do not seem concerning initially, except that the range of values is quite large, suggesting that we may have high variability in the model, further suggesting that our data might not explain the data well.

Looking at the empirical density (in the first plot), we see first a point-mass at zero in the density, reflecting the zero-inflated nature of the data, as well as a strong right skew. The former may be removed by filtering out the zero-valued observations (and this will be formalized in a later model), and the latter may be corrected by applying a variable transform for which we must re-define our model by adding a `variable_transform`. To supply an inverse transform, we must provide a list with a `transform` object and `inv_transform` object---functions which are their own inverses. The model will be estimated on the transformed data, and then be back-transformed using the inverse transformation.

```{r}
BHF_spec_trans <- specify(
  model = "BHF",
  domain_name = "county",
  response_name = "obs_biomass",
  auxiliary_variables = c("tcc", "evc"),
  variable_transform = list(
    transform = function (y) {y^(1/3)},
    inv_transform = function (y) {y^3}
  )
)
```

And we can then fit a model and check the posterior predictives:

```{r, cache = TRUE, include=FALSE}
BHF_fit_trans = fit.basal_spec(
  BHF_spec_trans,
  idaho_obs_full[idaho_obs_full$obs_biomass != 0,]
)
```

```{r}
check.basal_fit(
  BHF_fit_trans
)
```

All of these plots are much more reassuring. The emerical pdf and cdf both closely match the data with only slight deviations, and the ranges on mean and variance distributions are much lower.

The final step is to compute estimates, for which we need some new data. We could provide the initial data, but here we provide population data:

```{r}
estimate.basal_fit(
  BHF_fit_trans,
  idaho_aux_aggregate
)
```

The different types of statistics are pretty flexible. Furthermore, estimation can occur on arbitrary small areas. For example, though we used county as the domain when specifying the model, we could estimate on the level of vegetation type:

```{r}
idaho_aux_evt_agg = idaho_aux_full %>%
  group_by(evt) %>%
  reframe(tcc = mean(tcc, na.rm = T),
          evc = mean(evc, na.rm = T))

estimate.basal_fit(
  BHF_fit_trans,
  idaho_aux_evt_agg,
  domain = "evt",
  stat = c(mean = mean, se = sd)
)
```

By default, `estimate()` provides estimates of the mean and variance, however, the estimated statistics can be changed by changing the `stat` argument. For example, to construct credible intervals, we can use the quantile function. In thisc ase, since we are working with extreme quantiles, we increase the number of draws.

```{r}
estimate.basal_fit(
  BHF_fit_trans,
  idaho_aux_aggregate,
  stat = c(se = sd, 
           lower = function(x) {quantile(x, 0.025, na.rm = T)},
           upper = function(x) {quantile(x, 0.975, na.rm = T)}),
  ndraws = 5000
)
```

This allows us to make inferences like ``This model assigns 95% posterior probability to the event where the mean biomass in Ada county, Idaho is as above''. Although, it's important to remember that this model is on modified data, where we excluded zeroes.


\section{Zero Inflated Estimators/Two-Stage Models}

\section{Notes}

\subsection{Advanced modeling with `brms`.}

\subsection{Auto-Aggregation}

\subsection{Notes on `check()`}

\subsection{Notes on posterior prediction}
